A framework for mentored data science research

### 2019-12-12 Rmd/jse-2019-ms.Rmd
updates to methods
added markdown table
many updates

---
title: 'A framework for undergraduate data science research'
author: "Frederick J. Boehm and Bret M. Hanlon"
date: "8/20/2019"
params:
  lastmod: !r lubridate::now()
bibliography: jse.bib
output: 
  - github_document
  - bookdown::pdf_document2
  - bookdown::html_document2
---

>>>>

---
title: 'A framework for mentored data science research'
author: "Frederick J. Boehm and Bret M. Hanlon"
date: "8/20/2019"
params:
  lastmod: !r lubridate::now()
bibliography: jse.bib
geometry: "left=2.5cm,right=2.5cm,top=2.5cm,bottom=2.5cm"
output: 
  bookdown::pdf_book:
    includes:
      in_header: head.tex
---

### 2019-12-12 Rmd/methods.Rmd
many updates

@@

With the goal of promoting transparency in our research, we encouraged students to 
use `git` for version control of their code and documents and to share their code via
the website Github (https://github.com). One student also enrolled in Karl Broman's class on 
tools for reproducible research. This class features `git` and Github throughout 
its lectures and activities.

>>>>

```{r t1, echo = FALSE}
library(magrittr)
step <- c(rep("Question formulation", 3),
          rep("Data acquisition", 3), 
          rep("Data analysis and visualization", 2), 
          rep("Presentation and communication", 3)
)
details <- c("List ideas", "Prioritize ideas", "Check data availability",
             "Download tweets and other data", "Parse tweet JSON", "Organize data as times series",
             "Exploratory plots", "Summary statistics",
             "Interpret and summarize graphs and summary statistics", "Prepare written materials and figures", "Share findings as posters, slides, and reports"
             )
tibble::tibble(step, details) %>%
#  dplyr::group_by(step) %>%
  gt::gt() %>%
  gt::tab_style(
    style = list(
      gt::cell_fill(color = "lightcyan"),
      gt::cell_text(style = "italic")
      ),
    locations = gt::cells_data(
      columns = dplyr::vars(details),
      rows = step == "Question formulation"
      )
  ) %>%
    gt::tab_style(
    style = list(
      gt::cell_fill(color = "pink"),
      gt::cell_text(style = "italic")
      ),
    locations = gt::cells_data(
      columns = dplyr::vars(details),
      rows = step == "Data acquisition"
      )
  ) %>%
  gt::tab_header(
    title = gt::md("Framework Overview"), subtitle = "t"
  ) %>%
  gt::tab_style(
    style = list(
      gt::cell_fill(color = "yellow"),
      gt::cell_text(style = "italic")
      ),
    locations = gt::cells_data(
      columns = dplyr::vars(details),
      rows = step == "Data analysis and visualization"
      )
  ) %>%
  gt::as_latex()
```

@@

With the goal of promoting transparency in our research, we encouraged students to 
use `git` for version control of their code and documents and to share their code via
the website Github (https://github.com). One student also enrolled in Karl Broman's class on 
tools for reproducible research (https://kbroman.org/Tools4RR/). This class features `git` and Github throughout 
its lectures and activities.

### 2019-12-12 Rmd/results.Rmd
many updates

>>>>

Students benefited from our emphasis on the four central concepts, 
three from @nolan2010computing plus reproducible research skills. 
The research projects successfully drew on emerging areas of 
statistical computing, namely text analysis. They combined 
computational topics, including topic modeling and time series 
methods, with data analysis in the practice of statistics. Although 
we didn't formally measure them, our informal assessment indicates 
that students' computational reasoning skills increased over the 
duration of our projects. Students used a variety of computing tools 
and methods to arrive at a practical solution to a selected task. 
They became more skilled in computing with R and shell scripts and 
more fluid in their verbal explanations during our regular meetings [@r].

Our framework's emphasis on reproducible research skills is 
evidenced by the students' R package, `parseTweetFiles`, which is 
both version controlled with `git` and shared via Github.com.
